# Data sources

## U.S. Energy Information Administration (EIA)
This dataset was pulled from EIA.gov through an API. There is a package in R called "eia" that helps automate data pulls from the EIA.gov API calls. We used this package to pull monthly power generation data by fuel type for all of the states in the USA from the time period of 2001 - 2021.
```{r, include = F, message = F, warning = F}
library(tidyverse)
library(kableExtra)
eia_dat <- read.csv('data/raw/monthly_fuel_generation_by_state_full.csv')
```

### Column Descriptions
The columns in this dataset are listed below: 

* **thousand_megawatthours:** net power generation in thousand megawatt hours
* **date:** contains the month in which the power was generated. This column ranges from "2001-01-01" to "2021-08-01".
* **state:** contains the state in which the power was generated. Unique values include all 50 USA states
* **fuel:** contains the fuel type, with unique values listed below:

```{r}
unique(eia_dat$fuel)
```
The fuels are aggregated hierarchically as outlined below:

* **All Fuels:** includes the aggregation of power generation for coal, petroleum liquids, petroleum coke, natural gas, other gases, nuclear, conventional hydroelectric, hydro-electric pumped storage, other, and other renewables (total).
* **Other renewables (total):** includes the aggregation of power generation for wind, all utility-scale solar, geothermal, and biomass (total).
* **Biomass (total):** includes the aggregation of power generation for wood and wodd-derive fuels and other biomass.
* **All solar:** includes the aggregation of power generation for small-scale solar photovoltaic and all utility-scale solar. 

The column summaries of data types and statistics (for numeric columns) are shown below:
```{r}
summary(eia_dat)
```
The dataset has dimensions of:
```{r}
dim(eia_dat)
```
or 145239 rows and 4 columns. 

This dataset only included power generation data for fuel types that were actually produced in a state. Therefore, some fuel type generation values are missing from certain states. Additionally, some states only adopted certain fuel types at a specific time interval during the period of interest, so again the power generation data is missing for some of the time period. This will be discussed more in the [missing values chapter](https://edav-ccfall21-final-project-team-7.github.io/USCleanEnergyAnalysis/missing-values.html)

