# Results
## Renewable Fuel Adoption
The below chart explores the average percent of power generated in 2021 by each state that is produced by renewable energy sources. Renewable energy sources include wind, solar, geothermal, biomass, and hydroelectric power sources.

```{r, warning=F, echo=F}
library(tidyverse)
library(GGally)
library(lubridate)
library(forecast)
library(gridExtra)
library(cowplot)
library(RColorBrewer)
library(usmap)
library(patchwork)
df <- read.csv('data/clean/clean_monthly_fuel_generation.csv')
# df$fuel <- trimws(df$fuel) 
df$date <- as.Date(df$date)
# df[is.na(df$thousand_megawatthours),]$thousand_megawatthours <- 0
# df$fuel <- gsub(" ", "_", df$fuel)
# df$fuel <- gsub("-", "_", df$fuel)
#Update solar summation to be correct
# df <- df %>% pivot_wider(id_cols = c(date, state), names_from=c(fuel), values_from = thousand_megawatthours)
# df <- df %>% rowwise() %>% mutate(all_solar = sum(small_scale_solar_photovoltaic, all_utility_scale_solar, na.rm = TRUE))
# df <- df %>% pivot_longer(cols=names(df)[3:21],  names_to='fuel', values_to='thousand_megawatthours')
```

```{r, warning=F, echo=F, fig.height=8}
current <- df %>% filter(year(date) == 2020)
current_states <- current %>% group_by(state, fuel) %>% summarize(total_thousand_mwh = mean(thousand_megawatthours, na.rm=TRUE)) %>% pivot_wider(id_cols = c(state), names_from= fuel, values_from = total_thousand_mwh)
current_states[is.na(current_states)] <- 0
current_states <- current_states[names(current_states) %in% c('state', 'coal', 'wind','all_solar', 'petroleum_liquids','petroleum_coke','natural_gas', 'other_gases', 'nuclear', 'conventional_hydroelectric', 'other', 'other_renewables_(total)')]
current_states$total <- current_states$coal + current_states$petroleum_coke + current_states$petroleum_liquids + current_states$natural_gas + current_states$other_gases + current_states$nuclear + current_states$conventional_hydroelectric + current_states$other + current_states$`other_renewables_(total)`
current_states$renewables <- current_states$conventional_hydroelectric + current_states$`other_renewables_(total)`

current_states$coal <- current_states$coal/current_states$total
current_states$renewables <- current_states$renewables/current_states$total
current_states$all_solar <- current_states$all_solar/current_states$total
current_states$wind <- current_states$wind/current_states$total


# current_states_percent <- current_states %>% 
#   gather(variable, value, -state) %>% 
#   group_by(state) %>% 
#   mutate(percentage = value/sum(value, na.rm=TRUE)*100) %>% 
#   select(-value) %>% 
#   spread(variable, percentage)

plot_df <- current_states[names(current_states) %in% c('state', 'all_solar', 'wind', 'renewables')]
plot_df <- plot_df %>% pivot_longer(cols = !state, names_to = "fuel", values_to = "percent") %>% mutate(fuel = replace(fuel, fuel == "all_solar", "solar"))

plot_df$fuel <- fct_relevel(plot_df$fuel, c('renewables', 'wind', 'solar'))


ggplot(plot_df, aes(x = percent, y = fct_reorder2(state, fuel == "renewables", percent, .desc = FALSE),
       color = fuel)) +
  geom_point() +
  scale_x_continuous(labels = scales::percent) +
  ggtitle("2020 Average Renewable Percent of Total Power Generation") +
  xlab("Net Generation (%)") +
  ylab("State") + 
  theme_linedraw(base_size = 12) +
  theme(plot.title = element_text(hjust=.5))
options(repr.P.width=4,repr.P.height=10)

```

It is interesting to note that Vermont is already generating 100% of their power from renewable energy sources. Additionally, Delaware has the lower percentage of power generated from renewables with these sources comprising just over 2% of the average 2020 power.

Some states have set Renewable Portfolio Goals (RPS) that specify that utilities must sell a certain percentage of electricity that is generated from renewable sources. Delaware set a goal in 2005 to achieve achieve power generation that is 25% renewables by 2025 (https://www.ncsl.org/research/energy/renewable-portfolio-standards.aspx); however, they have only achieved 2% renewables in the current year. 

## Macro Trends
To explore macro trends in the energy sources utilized by the United States as a whole, the time series for each power source is plotted from 2001 - 2021. It is seen below that in 2001 coal was the predominant electricity source for America. Coal had a steady power generation until it began to decline around 2010. Natural gas took off at this time and by 2021 is the biggest producer of energy. Conventional hydroelectric was one of the first renewable energy resources to appear at significant levels in the electricity power generation for the US. Both solar and wind power have slowly increased in the amount of power they produce. 

```{r, echo=F, warning=F, fig.width = 10}
df  %>% 
  group_by(fuel, date) %>% 
  summarise(mil_mwhrs = sum(thousand_megawatthours, na.rm = TRUE)/1000) %>% 
  filter(fuel %in% c('all_solar', 'coal','natural_gas', 'wind', 'nuclear', 'conventional_hydroelectric')) %>%
  mutate(fuel = replace(fuel, fuel == "conventional_hydroelectric", "hydroelectric")) %>% 
  mutate(fuel = replace(fuel, fuel == "all_solar", "solar")) %>% 
  ggplot(aes(x = date, y = mil_mwhrs, color = fct_relevel(fuel, c('natural_gas', 'coal','nuclear', 'hydroelectric', 'wind', 'solar')))) +
  geom_line() +
  ggtitle('USA Net Generation by Fuel Type') + 
  xlab("Date") +
  ylab("MW Hours (Mil)") +
  theme_classic(12)+
  theme(legend.title=element_blank(), plot.title = element_text(hjust=.5))
```

We can explore the relationship between specific fuel types to determine which are correlated. From the plot below, it appears that coal is negatively correlated with both solar and natural gas. This is not necessarily surprising because natural gas is typically used to complement renewable energy sources and act as a form of stability to produce power when the sun isn't shining or the wind isn't blowing. 

```{r, echo=F, warning=F}
# df  %>% filter(fuel %in% c('wind','coal', 'natural_gas', 'all_solar')) %>%
#   select(c(fuel, date, thousand_megawatthours)) %>% 
#   group_by(fuel, date) %>% 
#   summarise(thousand_megawatthours = sum(thousand_megawatthours, na.rm = TRUE)) %>% ungroup()  %>%
#   pivot_wider(names_from=fuel, id_cols = date, values_from = thousand_megawatthours) %>%
#   ggparcoord(columns=c(array(2:5)), scale= "uniminmax", alphaLines = .3, splineFactor = 9) +
#   theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
#         plot.title = element_text(hjust=.5)) +
#   ggtitle('Relationship between Fuel Types') +
#   xlab("Fuel") +
#   ylab("") 

```

We can explore the relationship between specific fuel types to determine which are correlated. From the plot below, we can see that solar is negatively correlated with coal and positively correlated with natural gas and wind. This is not necessarily surprising because natural gas is typically used to complement variable renewable energy sources like solar and wind and act as a form of stability to produce power when the sun isn't shining or the wind isn't blowing. It makes sense that as the adoption of wind and solar energy increases, natural gas usage will also increase until a cleaner alternative is found. Natural gas is a cleaner fuel source than coal and produces 50% less CO\_2 emissions (https://group.met.com/fyouture/natural-gas-vs-coal/66). However, the cleanest alternative would be another renewable energy source that could help address the variability in production of wind and solar, like batteries.  

```{r, echo=F, warning=F}
df  %>% filter(fuel %in% c('wind','coal', 'natural_gas', 'all_solar', 'nuclear')) %>%
  group_by(fuel, date) %>% summarise(million_mwhrs = sum(thousand_megawatthours, na.rm = TRUE)/1000) %>%
  pivot_wider(names_from=fuel, id_cols = c(date), values_from = million_mwhrs) %>% ungroup() %>% select(-c( date)) %>% ggpairs(lower=list(continuous=wrap(ggally_smooth, color='darkgray', size=1)),
  diag=list(continuous=wrap(ggally_barDiag, color='black')), 
  upper=list(axisLabels='show',continuous=wrap('cor', color='black'))) + 
  xlab('MW Hours (Mil)') +
  ylab('MW Hours (Mil)') +
  theme_bw() +
  theme(legend.position = "none", 
        panel.grid.major = element_blank(), 
        axis.ticks = element_blank(), 
        panel.border = element_rect( colour = "black", fill = NA),
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
        plot.title = element_text(hjust=.5)) +
  ggtitle('Correlation of Fuel Source Power Generation')
```

To explore the seasonal component, look at the seasonplot for 2015-2021 when solar was adopted more widely as a fuel type. The solar power net generation seasonal trend peaks in the summer months (May - August) and dip in the winter months. This is most likely due to the fact that there is more sunshine in the summer months, so more solar power is able to be generated. The coal plot shows the power generation trends overall. There is a peak in June - September and December - February. These peaks make sense because these are the times of year that energy demand is at the highest due to heating demand during the cold months and cooling demand during the hot months. In the seasonal plot, we can see that solar is increasing in production from 2015-201, while coal is decreasing in production.

```{r, warning=F, echo =F}
my_colors <- RColorBrewer::brewer.pal(9, "Greens")[3:9]
seasonal_plot_data <- df  %>% filter(fuel %in% c('wind','coal', 'natural_gas', 'all_solar', 'nuclear')) %>%
  group_by(fuel, date) %>% summarise(thousand_megawatthours = sum(thousand_megawatthours, na.rm = TRUE)) %>%
  pivot_wider(names_from=fuel, id_cols = c(date), values_from = thousand_megawatthours) %>% ungroup()
seasonal_plot_data <- seasonal_plot_data %>% mutate(month_val = month(date))
seasonal_plot_data$date <- as.Date(seasonal_plot_data$date)

seasonal_plot_data <- seasonal_plot_data %>% filter(year(seasonal_plot_data$date) > 2014)
solar_series <- ts(seasonal_plot_data$all_solar/1000, start = c(2015,1,1), frequency = 12)
wind_series <- ts(seasonal_plot_data$wind/1000, start = c(2015,1,1), frequency = 12)
coal_series <- ts(seasonal_plot_data$coal/1000, start = c(2015,1,1), frequency = 12)
g1 <- ggseasonplot(solar_series) +
  labs(title = "Solar") +
  xlab("Month") + 
  ylab("MW Hours (Mil)") +
  guides(colour = guide_legend(reverse = TRUE)) +
  theme_bw(12) +
  theme(plot.title = element_text(hjust = 0.5)) + 
  scale_color_manual(values=my_colors) +
  theme(legend.position = "none")

g2 <- ggseasonplot(coal_series) +
  labs(title = "Coal") +
  xlab("Month") + 
  ylab("MW Hours (Mil)") +
  guides(colour = guide_legend(reverse = TRUE)) +
  theme_bw(12) +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(values=my_colors) +
  theme(legend.position = "none")

gplaceholder <- ggseasonplot(coal_series) +
  guides(colour = guide_legend(reverse = TRUE)) +
  theme_bw(12) +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(values=my_colors) +
  theme(legend.position = "right")

extract_legend <- function(my_ggp) {
  step1 <- ggplot_gtable(ggplot_build(my_ggp))
  step2 <- which(sapply(step1$grobs, function(x) x$name) == "guide-box")
  step3 <- step1$grobs[[step2]]
  return(step3)
}

# Apply user-defined function to extract legend
shared_legend <- extract_legend(gplaceholder)

#grid.arrange(arrangeGrob(g1,shared_legend,g2, nrow=2, ncol=2) ,widths = c(10, 1), top=textGrob("Seasonal #Plot for Total Power Generation", gp=gpar(fontsize=16,font=8)))

plot_row <- plot_grid(g1,shared_legend, g2, align = "v", nrow = 2, rel_widths  = c(7/8, 1/8, 7/8))

title <- ggdraw() + 
  draw_label(
    "Seasonal Plot for Total Power Generation by Fuel",
    x = .5,
    hjust = .5
  ) +
  theme(
    # add margin on the left of the drawing canvas,
    # so title is aligned with left edge of first plot
    plot.margin = margin(0, 0, 0, 7)
  )
plot_grid(
  title, plot_row,
  ncol = 1,
  # rel_heights values control vertical title margins
  rel_heights = c(0.1, 1)
)
#https://wilkelab.org/cowplot/articles/plot_grid.html - cowplot
#shared legend function: https://statisticsglobe.com/add-common-legend-to-combined-ggplot2-plots-in-r/

```

The most noticeable change in primary fuels can be seen from coal and natural gas. In 2001, 32 states used coal as the primary fuel, while by 2020 this number changed to 11. Natural gas experienced the opposite trend where it grew from being the primary fuel for 7 states to 22. nuclear, hydroelectric, and petroleum liquids were also primary fuels for states in 2001 and 2020. These fuels did not see much change in number of states using them as their primary fuel. Renewables (other than hydroelectric) did become the primary fuel for 4 states (Iowa, Kansas, Maine, and Minnesota) in 2020, while in 2001 no states were using other renewables as the primary fuel.

```{r, echo=F, warning=F}
max_fuel_2020 <- df %>% filter(year(date) == 2020) %>%
  select(c(fuel, state, thousand_megawatthours)) %>% 
  group_by(fuel, state) %>% 
  summarise(thousand_megawatthours = sum(thousand_megawatthours, na.rm = TRUE)) %>% ungroup()  %>%
  group_by(state) %>% filter(thousand_megawatthours == max(thousand_megawatthours)) 

max_fuel_2020 <- max_fuel_2020 %>% group_by(fuel) %>% summarize(count=n())
max_fuel_2020$year <- '2020'

max_fuel_2001 <- df %>% filter(year(date) == 2001) %>%
  select(c(fuel, state, thousand_megawatthours)) %>% 
  group_by(fuel, state) %>% 
  summarise(thousand_megawatthours = sum(thousand_megawatthours, na.rm = TRUE)) %>% ungroup()  %>%
  group_by(state) %>% filter(thousand_megawatthours == max(thousand_megawatthours)) 

max_fuel_2001 <- max_fuel_2001 %>% group_by(fuel) %>% summarize(count=n())
max_fuel_2001$year <- '2001'

full_df <- rbind(max_fuel_2020, max_fuel_2001)
full_df$fuel[full_df$fuel == "conventional_hydroelectric"] <- 'hydroelectric'
full_df$fuel[full_df$fuel == "other_renewables_(total)"] <- 'other_renewables'
fuel <- 'other_renewables'
count <- 0
year <- '2001'
missing_row <- data.frame(fuel, count,year)
full_df <- rbind(full_df, missing_row)

ggplot(full_df, aes(x = fct_reorder2(fuel, year == '2001', count), y = count, fill =year)) +
  geom_col(position = "dodge") +
  xlab("Primary Fuel") +
  ylab("State Count") +
  ggtitle("Count of States' Primary Fuel Source") +
  theme_classic() +
  theme(plot.title = element_text(hjust=.5), axis.text.x = element_text(angle = 45, hjust=1)) + scale_fill_brewer(palette = "Paired")

```

The heatmaps below provide insights into how the energy production is distributed over the United States. It can be noticed that California is the major producer of solar energy whereas Texas is the major producer of Coal, Natural Gas, and Wind energy. 

```{r,fig.width=16,fig.height=16}
df_all_solar <- df  %>% 
  group_by(fuel, date, state) %>% summarise(thousand_megawatthours = sum(thousand_megawatthours)) %>% filter(fuel %in% c('all_solar'))

df_coal <- df  %>% 
  group_by(fuel, date, state) %>% summarise(thousand_megawatthours = sum(thousand_megawatthours)) %>% filter(fuel %in% c('coal'))

df_natural_gas <- df  %>% 
  group_by(fuel, date, state) %>% summarise(thousand_megawatthours = sum(thousand_megawatthours)) %>% filter(fuel %in% c('natural_gas'))

df_wind <- df  %>% 
  group_by(fuel, date, state) %>% summarise(thousand_megawatthours = sum(thousand_megawatthours)) %>% filter(fuel %in% c('wind'))

df_all_solar1 <- df_all_solar %>% group_by(state) %>% summarize(thousand_megawatthours = mean(thousand_megawatthours)) 

df_coal1 <- df_coal %>% group_by(state) %>% summarize(thousand_megawatthours = mean(thousand_megawatthours)) 

df_natural_gas1 <- df_natural_gas %>% group_by(state) %>% summarize(thousand_megawatthours = mean(thousand_megawatthours)) 

df_wind1 <- df_wind %>% group_by(state) %>% summarize(thousand_megawatthours = mean(thousand_megawatthours)) 



g1 <- plot_usmap(data = df_all_solar1, values = 'thousand_megawatthours',regions = 'states', labels = TRUE, color = 'pink')+
  scale_fill_continuous(low = "pink", high = "red", name = "Average thousand megawatthours", label = scales::comma) + 
  labs(title = "Average thousand megawatthours by State (Solar Energy)") +
  theme(legend.position = "right",plot.title = element_text(hjust = 0.5,size=16))

g2 <- plot_usmap(data = df_coal1, values = 'thousand_megawatthours',regions = 'states', labels = TRUE, color = 'pink')+
  scale_fill_continuous(low = "pink", high = "red", name = "Average thousand megawatthours", label = scales::comma) + 
  labs(title = "Average thousand megawatthours by State (Coal)") +
  theme(legend.position = "right",plot.title = element_text(hjust = 0.5,size=16))

g3 <- plot_usmap(data = df_natural_gas1, values = 'thousand_megawatthours',regions = 'states', labels = TRUE, color = 'pink')+
  scale_fill_continuous(low = "pink", high = "red", name = "Average thousand megawatthours", label = scales::comma) + 
  labs(title = "Average thousand megawatthours by State (Natual Gas)") +
  theme(legend.position = "right",plot.title = element_text(hjust = 0.5,size=16))

g4 <- plot_usmap(data = df_wind1, values = 'thousand_megawatthours',regions = 'states', labels = TRUE, color = 'pink')+
  scale_fill_continuous(low = "pink", high = "red", name = "Normalized Average thousand megawatthours", label = scales::comma) + 
  labs(title = "Average thousand megawatthours by State (Wind Energy)") +
  theme(legend.position = "right",plot.title = element_text(hjust = 0.5,size=16))



g1+g2+g3+g4+plot_layout(ncol=2,nrow=2,widths=c(8,8),heights=c(38,38))+plot_annotation(title = "Raw values for energy produced",theme = theme(plot.title = element_text(size = 36,face='bold')))
```

However, we know that the size of every state is the United States varies a lot. Therefore, to account for the size and provide a better comparison of the energy produced in a state, we normalize the amount of energy produced by the area of the state and plot it on a heatmap as done above. It can be noticed from this heatmap that Texas is no longer one of the dominating producers of coal and natural gas. 

```{r,fig.width=16,fig.height=16}
area <- as.data.frame(read_csv('data/raw/area.csv'))

df_all_solar2 <- merge(df_all_solar1,area,by='state')
df_all_solar2$thousand_megawatthours<-df_all_solar2$thousand_megawatthours/df_all_solar2$area

df_coal2 <- merge(df_coal1,area,by='state')
df_coal2$thousand_megawatthours<-df_coal2$thousand_megawatthours/df_coal2$area

df_natural_gas2 <- merge(df_natural_gas1,area,by='state')
df_natural_gas2$thousand_megawatthours<-df_natural_gas2$thousand_megawatthours/df_natural_gas2$area

df_wind2 <- merge(df_wind1,area,by='state')
df_wind2$thousand_megawatthours<-df_wind2$thousand_megawatthours/df_wind2$area


g1 <- plot_usmap(data = df_all_solar2, values = 'thousand_megawatthours',regions = 'states', labels = TRUE, color = 'pink')+
  scale_fill_continuous(low = "pink", high = "red", name = "Normalized Average thousand megawatthours", label = scales::comma) + 
  labs(title = "Normalized Average thousand megawatthours by State (Solar Energy)") +
  theme(legend.position = "right",plot.title = element_text(hjust = 0.5,size=16))

g2 <- plot_usmap(data = df_coal2, values = 'thousand_megawatthours',regions = 'states', labels = TRUE, color = 'pink')+
  scale_fill_continuous(low = "pink", high = "red", name ="Normalized Average thousand megawatthours", label = scales::comma) + 
  labs(title = "Normalized Average thousand megawatthours by State (Coal)") +
  theme(legend.position = "right",plot.title = element_text(hjust = 0.5,size=16))

g3 <- plot_usmap(data = df_natural_gas2, values = 'thousand_megawatthours',regions = 'states', labels = TRUE, color = 'pink')+
  scale_fill_continuous(low = "pink", high = "red", name = "Normalized Average thousand megawatthours", label = scales::comma) + 
  labs(title = "Normalized Average thousand megawatthours by State (Natual Gas)") +
  theme(legend.position = "right",plot.title = element_text(hjust = 0.5,size=16))

g4 <- plot_usmap(data = df_wind2, values = 'thousand_megawatthours',regions = 'states', labels = TRUE, color = 'pink')+
  scale_fill_continuous(low = "pink", high = "red", name = "Normalized Average thousand megawatthours", label = scales::comma) + 
  labs(title = "Normalized Average thousand megawatthours by State (Wind Energy)") +
  theme(legend.position = "right",plot.title = element_text(hjust = 0.5,size=16))


g1+g2+g3+g4+plot_layout(ncol=2,nrow=2,widths=c(8,8),heights=c(38,38))+plot_annotation(title = "Energy Produced per sq miles (by state)",theme = theme(plot.title = element_text(size = 36,face='bold')))

```

However, it should be kept in mind that the actual energy production is represented by the first plot where california and texas are dominating producers of energy and the second plot only slows the energy produced per sq miles for every state. 

## Correlation of Energy Consumption, CO2 Emission and Economic Indicators

```{r, echo=F, warning=F}

library(tidyverse)
library (dplyr)
library(GGally)

us_wb_df <- as.data.frame(read_csv('data/raw/clean_worldbank.csv'))
row.names(us_wb_df)<-us_wb_df[,"Series"]
us_wb_df <-us_wb_df[,2:length(names(us_wb_df))]

rownames(us_wb_df) <- c('Pop T', 'Pop G', 'Area', 'Pov Rt', 'GNI T','GNI Pc', 'GNI PPP', 'GNI PPP Pc', 'Inc Shr', 'Life Exp', 'Fertility Rt', 'Ado Ft', 'Contr Pr', 'Birth Rt', 'Mort Rt', 'Under Wt', 'Imm Mea', 'Prim Comp', 'Sec Sch', 'Prim Sch', 'HIV', 'Forest A', 'Water P', 'Energy Use', 'C02 Em', 'Elec Pwr', 'GDP T', 'GDP %', 'Infl %', 'Agri %', 'Industry %', 'Exports %', 'Imports %', 'Gross %', 'Revenue, %', 'Start up', 'Mkt Cap %', 'Military %', 'Mobile Sub', 'HT Exp %', 'Merchandise %', 'Net Bart', 'Ext Debt', 'Debt Sv %', 'Migration', 'Remittance', 'FDI', 'ODA')

my_us_wb_df_p <- as.data.frame(t(us_wb_df))

my_us_wb_df_p <- my_us_wb_df_p %>% relocate('GDP %', .before = 'GDP T')
my_us_wb_df_p <- my_us_wb_df_p %>% relocate('GNI Pc', .before = 'GDP %')

ggparcoord(my_us_wb_df_p, columns = 23:27, scale = "globalminmax") + 
  geom_vline(xintercept = 1:8, color = "lightblue") +
  ggtitle("US Macro Energy Consumption, CO2 Emission, & Economic Indicators ") + 
  xlab("") + 
  ylab("") +
  theme(plot.title = element_text(hjust=.5))


ggparcoord(my_us_wb_df_p, columns = 23:27, scale = "uniminmax", alphaLines = 0.3, splineFactor = 10) + 
  geom_vline(xintercept = 1:8, color = "lightblue") +
  ggtitle("US Macro Energy Consumption, CO2 Emission, & Economic Indicators ") + 
  xlab("") + 
  ylab("") +
  theme(plot.title = element_text(hjust=.5))


```


## Correlation between Energy ETF and macro trends in energy
```{r}
min_max_norm <- function(x) {
    (x - min(x)) / (max(x) - min(x))
  }

oil <- read_csv('data/clean/clean_crushing_crude_oil.csv')
oil$Day <- as.Date(oil$Day,format = '%m/%d/%Y')
names(oil) <- c("Date","Cost")


#cnrg <-read_csv('data/raw/CNRG.csv')
#icln <- read_csv('data/raw/ICLN.csv')
xle <- read_csv('data/raw/XLE.csv')



#cnrg$Close<-as.data.frame(lapply(cnrg[5],min_max_norm))$Close
#icln$Close<-as.data.frame(lapply(icln[5],min_max_norm))$Close
xle$Close<-as.data.frame(lapply(xle[5],min_max_norm))$Close


energy_df <- df  %>% 
  group_by(fuel, date) %>%
  summarise(thousand_megawatthours = sum(thousand_megawatthours, na.rm = TRUE)) %>%
  filter(fuel %in% c('all_solar', 'coal','natural_gas', 'wind', 'nuclear', 'conventional_hydroelectric'))

total_energy <- energy_df %>% 
  group_by(date) %>% 
  summarise(thousand_megawatthours = sum(thousand_megawatthours, na.rm = TRUE))

total_energy$thousand_megawatthours<-as.data.frame(lapply(total_energy[2],min_max_norm))$thousand_megawatthours

names(total_energy) <- c('Date','thousand_megawatthours')

colors<-c("Total enegry (normalized)" = "red", "Normalized Closing Price (XLE)" = "blue")
merged_data <-merge(xle,total_energy,by='Date')

g1<-ggplot(merged_data)+
  geom_line(aes(Date,thousand_megawatthours,color="Total enegry (normalized)"))+
  geom_line(aes(Date,Close,color="Normalized Closing Price (XLE)"))+
  labs(x='Year', y='normalized scale',color='Legend')


df_solar2 <- df_all_solar %>% 
  group_by(date) %>% 
  summarize(solar_megawatt = sum(thousand_megawatthours)) 

df_solar2$solar_megawatt<-as.data.frame(lapply(df_solar2[2],min_max_norm))$solar_megawatt

df_coal2 <- df_coal %>% 
  group_by(date) %>% 
  summarize(coal_megawatt = sum(thousand_megawatthours)) 

df_coal2$coal_megawatt<-as.data.frame(lapply(df_coal2[2],min_max_norm))$coal_megawatt

df_natural_gas2 <- df_natural_gas %>% 
  group_by(date) %>% 
  summarize(gas_megawatt = sum(thousand_megawatthours)) 

df_natural_gas2$gas_megawatt<-as.data.frame(lapply(df_natural_gas2[2],min_max_norm))$gas_megawatt

df_wind2 <- df_wind %>% 
  group_by(date) %>% 
  summarize(wind_megawatt = sum(thousand_megawatthours)) 

df_wind2$wind_megawatt<-as.data.frame(lapply(df_wind2[2],min_max_norm))$wind_megawatt

names(xle)<-c("date","Open","High","Low","Close","Adj Close","Volume")

colors<-c("solar"="goldenrod3","coal"="grey1","natural gas"="darkolivegreen3","wind"="dodgerblue3","close price (XLE)"="red")

merged_data1 <-merge(xle,df_solar2,by='date')
merged_data2 <-merge(merged_data1,df_coal2,by='date')
merged_data3 <-merge(merged_data2,df_natural_gas2,by='date')
merged_data <-merge(merged_data3,df_wind2,by='date')

g2<-ggplot(merged_data)+
  geom_line(aes(date,solar_megawatt,color="solar"))+
  geom_line(aes(date,coal_megawatt,color="coal"))+
  geom_line(aes(date,gas_megawatt,color="natural gas"))+
  geom_line(aes(date,wind_megawatt,color="wind"))+
  geom_line(aes(date,Close,color="close price (XLE)"))+
  labs(x='Year', y='normalized scale',color='Legend')+
  scale_color_manual(values=colors)


g1+g2+plot_layout(nrow=2,ncol=1)

```


```{r}
df1 <- df  %>% 
  group_by(fuel, date) %>% 
  summarise(mil_mwhrs = sum(thousand_megawatthours, na.rm = TRUE)/1000) %>% 
  filter(fuel %in% c('all_solar', 'coal','natural_gas', 'wind', 'nuclear', 'conventional_hydroelectric')) %>%
  mutate(fuel = replace(fuel, fuel == "conventional_hydroelectric", "hydroelectric")) %>% 
  mutate(fuel = replace(fuel, fuel == "all_solar", "solar"))


df2<-merge(df1,xle,by='date')
df2$Close

  ggplot(df2) +
  geom_line(aes(x = date, y = mil_mwhrs,color = fct_relevel(fuel, c('natural_gas', 'coal','nuclear', 'hydroelectric', 'wind', 'solar')))) +
  geom_line(aes(x=date,y=Close*175))+
  ggtitle('USA Net Generation by Fuel Type') + 
  xlab("Date") +
  ylab("MW Hours (Mil)") +
  theme_classic(12)+
  theme(legend.title=element_blank(), plot.title = element_text(hjust=.5))
```

