---
title: "eia_data_analysis"
author: "Gabrielle Nyirjesy"
date: "11/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(data.table)
library(d3r)
library(parcoords)
library(GGally)
library(janitor)
library(egg)
library(ggpubr)
```
```{r}
find_missing_patterns <- function(df, aggregation = 'count'){
  
  #find missing patterns
  missing_patterns <- data.frame(is.na(df)) %>%
  group_by_all() %>%
  count(name = "count", sort = TRUE) %>%
  ungroup()
    #create g1
  missing_patterns_sums <- df %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>% 
  pivot_longer(cols = names(df), names_to='name',values_to='value') %>%
  arrange(desc(value))
  
   if (aggregation=='percent'){
    total_sum<-length(row.names(df))
    missing_patterns_sums$value<- missing_patterns_sums$value*100/total_sum
    print(total_sum)
    y_label = 'percent rows \n missing'
   }
  else{
    y_label = 'num rows \n missing'
  }
  column_reorder <- missing_patterns_sums$name
  
  
  g1 <- ggplot(missing_patterns_sums, aes(x=fct_reorder(name, -value), y=value))+
    geom_bar(stat='identity', fill='#9eb7ee') +
    scale_y_continuous(limits=c(0, max(missing_patterns_sums$value)), expand=c(0,0)) +
    xlab('') +
    ylab(y_label) +
    ggtitle('Missing value patterns') +
    theme(panel.border = element_rect(colour = "black", fill=NA, size=.5), 
          panel.background = element_rect(fill = "white"),
          panel.grid.minor.x = element_blank(),
          panel.grid.minor.y = element_blank(),
          panel.grid.major.x = element_blank(),
          panel.grid.major.y = element_line(color='darkgray', size=.1)) +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
    theme(aspect.ratio = .25)
  
  #Create g3
  missing_patterns_count <- missing_patterns %>% mutate_all((as.integer)) %>% adorn_totals("col")
  missing_patterns_count$col <- as.factor(ifelse(missing_patterns_count$Total - missing_patterns_count$count == 0 , 1, 0.5))
  missing_patterns_count$id <- row.names(missing_patterns_count)
  
  if(aggregation=='percent'){
    missing_patterns_count$count<-missing_patterns_count$count*100/sum(missing_patterns_count$count)
    x_label = 'percent rows'
  }
  else{
    x_label = 'row count'
  }
  pattern_id <- missing_patterns_count  %>% filter(col == 1) %>% select(id) %>% as.character()
  missing_patterns_count$id <- factor(missing_patterns_count$id)
  missing_patterns_count$id <- fct_relevel(missing_patterns_count$id, paste(sort(as.integer(levels(missing_patterns_count$id)), decreasing = TRUE)))
  g3 <- missing_patterns_count %>%
    #rowid_to_column(var="pattern") %>%
    #mutate(pattern = factor(pattern)) %>%
    ggplot(aes(y=fct_reorder(id, -desc(count)), x=count)) + 
    geom_bar(stat='identity', fill='#9eb7ee',aes(alpha = factor(missing_patterns_count$col))) +
    scale_x_continuous(limits=c(0, max(missing_patterns_count$count)), expand=c(0,0)) + #, breaks = seq(0,max(missing_patterns$count), 5) if want to have the y axis break on 5s
    ylab('') +
    xlab(y_label) +
    theme(panel.border = element_rect(colour = "black", fill=NA, size=.5), 
          panel.background = element_rect(fill = "white"),
          panel.grid.minor.x = element_blank(),
          panel.grid.minor.y = element_blank(),
          panel.grid.major.y = element_blank(),
          panel.grid.major.x = element_line(color='darkgray', size=.1),) +
    scale_alpha_manual(values=c("0.5"=0.5,"1"=1),guide="none") +
    theme(aspect.ratio = 4)
  
  #create g2
  main_plot_df <- missing_patterns %>%
  select(-c(count)) %>%
  as_tibble() %>%
  rowid_to_column(var="Y") %>%
  gather(key="X", value="Z", -1) 

  main_plot_df$X <- as.factor(main_plot_df$X)
  main_plot_df$Y <- as.factor(main_plot_df$Y)
  main_plot_df$X <- fct_relevel(main_plot_df$X, column_reorder)
  main_plot_df$Y <- fct_relevel(main_plot_df$Y, rev(levels(main_plot_df$Y)))
  
  
  main_plot_df$Z <- as.factor(ifelse(main_plot_df$Y == pattern_id, .5, main_plot_df$Z))
  
  
  g2 <- ggplot(main_plot_df, aes(X, Y, fill= Z)) + 
      geom_tile(color='white') +
      theme(legend.position="none") +
    scale_fill_manual(values=c('#cacaca', 'darkgray', '#b2a0e1')) +
    annotate("text",x=length(unique(main_plot_df$X))/2,y=length(unique(main_plot_df$Y))-as.integer(pattern_id)+1,label='complete cases') + 
    #xlab('Variable') +
    ylab('Missing Pattern') +
    theme_classic() +
    theme(legend.position = "none") +
    theme(axis.text.x = element_text(angle = 45, vjust =1, hjust=1)) +
    coord_equal() + theme(aspect.ratio = length(row.names(missing_patterns))/(length(names(missing_patterns))-1))
  
  #Arrange graphs
  blank <- ggplot() + 
              geom_point(aes(1,1), colour="white") +
              theme(                              
                plot.background = element_blank(), 
                panel.grid.major = element_blank(), 
                panel.grid.minor = element_blank(), 
                panel.border = element_blank(), 
                panel.background = element_blank(),
                axis.title.x = element_blank(),
                axis.title.y = element_blank(),
                axis.text.x = element_blank(),
                axis.text.y = element_blank(),
                axis.ticks = element_blank()
              )

  ggarrange(g1,blank,g2,g3,ncol=2,nrow=2,widths = c(16,4),heights=c(1,4))
  #return(missing_patterns_count)
}
```


```{r}
#https://www.eia.gov/electricity/data/browser/#/topic/0?freq=M&start=200101&end=202108&ctype=linechart&ltype=pin&pin=&maptype=0
df <- read.csv('data/raw/monthly_fuel_generation_by_state_full.csv')
df$fuel <- trimws(df$fuel) 
df$date <- as.Date(df$date)
df[is.na(df$thousand_megawatthours),]$thousand_megawatthours <- 0
df$fuel <- gsub(" ", "_", df$fuel)
df$fuel <- gsub("-", "_", df$fuel)
```

way to view the n/a columns for dates
```{r, fig.height = 10, fig.width = 10}
test <- df %>% pivot_wider(id_cols = c(state, fuel), names_from= date, values_from = thousand_megawatthours)
test
```
View which fuel types are missing for each state
```{r, fig.height = 10, fig.width = 10}
test2 <- df %>% group_by(state, fuel) %>% summarize(total_thousand_mwh = sum(thousand_megawatthours)) %>% pivot_wider(id_cols = c(state), names_from= fuel, values_from = total_thousand_mwh)
test2
test3 <- test2
test3$state <- NULL
find_missing_patterns(test3, 'percent')
#ggsave('state-fuel type-missing patterns.png', height = 10, width = 10, units = "in")
```


```{r}
df %>% filter(state == 'USA-AL') %>% 
  ggplot(aes(x = date, y = thousand_megawatthours, color = fuel)) + 
  geom_line() #+ 
  #facet_wrap(~fuel)#, ncol = 1, scales="free")

#could do a boxplot for each state, or even a cleveland dot plot
```
```{r}
df %>% filter(state == 'USA-CA') %>% 
  group_by(fuel) %>% mutate(thousand_megawatthours_norm = scale(thousand_megawatthours)) %>%
  ggplot(aes(x = date, y = thousand_megawatthours_norm)) +
  facet_wrap(~fuel) +
  geom_line()
```
```{r}
df  %>% 
  group_by(fuel, date) %>% summarise(thousand_megawatthours = sum(thousand_megawatthours)) %>% filter(fuel %in% c('all_solar', 'all_utility_scale_solar', 'coal','natural_gas', 'wind')) %>%
  ggplot(aes(x = date, y = thousand_megawatthours, color = fuel)) +

  geom_line()
```
Missing values: "all solar" values don't begin until 2015; however, "all utility scale solar" has values beginning in 2001. The aggregation is off here for the all_utility_scale_solar section
```{r}

df %>% group_by(fuel, date) %>% 
  summarise(thousand_megawatthours = sum(thousand_megawatthours)) %>% 
  filter(fuel %like% 'solar') %>%
  ggplot(aes(x = date, y = thousand_megawatthours, color = fuel)) +
  geom_line()
```
for the plots to make sense, need to set the all solar value to be all_utility_scale_solar for the time period before small_scale_solar_photovoltaic had values
```{r}
#df$thousand_megawatthours[df$]
```


look at which fuel sources are correlated
```{r}
df %>% select(c(fuel, date, thousand_megawatthours)) %>% 
  group_by(fuel, date) %>% 
  summarise(thousand_megawatthours = sum(thousand_megawatthours)) %>% ungroup() %>%
  pivot_wider(names_from=fuel, id_cols = date, values_from = thousand_megawatthours) %>% 
  ggplot(aes(x=all_utility_scale_solar, y = natural_gas)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  ggtitle("Natural Gas vs. Solar") +
  theme(plot.title = element_text(hjust = .5),
          panel.grid.major = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"))
```
plot of percent growth for each fuel type over the entire time period - could do this as a cleveland dot plot - should probably make the values annualized before doing the growth values becuase the numbers are seasonal
```{r}
df %>% select(c(fuel, date, thousand_megawatthours)) %>% 
  group_by(fuel, date) %>% 
  summarise(thousand_megawatthours = sum(thousand_megawatthours)) %>% ungroup() %>%
  pivot_wider(names_from=fuel, id_cols = date, values_from = thousand_megawatthours) %>%
  ggparcoord(columns=c(array(2:20)), scale= "uniminmax", alphaLines = .3, splineFactor = 9) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

```
```{r}
comp_df <-df %>% select(c(state, fuel, date, thousand_megawatthours)) %>% 
  group_by(state, fuel, date) %>% 
  summarise(thousand_megawatthours = sum(thousand_megawatthours)) %>% ungroup() %>%
  pivot_wider(names_from=fuel, id_cols = c(date, state), values_from = thousand_megawatthours)

  ggparcoord(comp_df, columns=c(array(3:21)), scale= "uniminmax", alphaLines = .3, splineFactor = 9, groupColumn = "state") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

Checking correlation between financial data and amount of energy produced
```{r}

#Why is hydro negative?

unique(df$fuel)
```

Strip off USA- from state col and plot them on a map


Missing data plots
```{r}

WorldBank <-read_csv('data/raw/US_Country_Data_WorldBank.csv')

png(file='test1.jpeg')
find_missing_patterns(WorldBank,'percent')
dev.off()
```

